### 2014-11-13 evaluation.tex
edits

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under a 1DC configuration. The YCSB workload was modified to issue 10
operations per transaction, with a default 50:50 read:write mix. Each operation
is assumed to be eventually consistent. \cf{NoTxn} corresponds to a
configuration that does not use transactions. Compared to this, RC is only 12\%
shower in terms of latency with 512 clients, whereas RR is 2.3X slower. The
difference between RC and \cf{NoTxn} is due to the meta-data overhead of
recording transaction information in the object state. For RR transactions, the
cost of capturing and maintaining a snapshot is the biggest source of overhead.

@@

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
configuration against a single replica ({\sf\small NoRep}) and a strongly
replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
default RUBiS bidding mix, which has 15\% read-write interactions, which is
representative of the auction workload.  Without replication, {\sf\small NoRep}
trivially provides strong consistency. However, this deployment does not scale
beyond 1750 operations per second. Strong replication offers better throughput
at the cost of greater latency due to inter-replica coordination. The \name
deployment offers the benefit of replication, while only paying the cost of
coordination when necessary.

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We use 128 clients and a single \name replica,
with all clients operating on the \emph{same} LWW register to stress test the
summarization mechanism. The shim layer cache (memory) is summarized every 64
updates, while the updates in the backing store (disk) are summarized every
4096 updates. Each point in the graph represents the average latency of the
previous 1000 operations. Each experiment is run for one minute.  Without
summarization, the average latency of operations increases exponentially to
almost one second, and only 13K operations were performed in a minute. Since
every operation has to reduce over the set of all previous operations,
operations take increasingly more time to complete since they must contend with
an ever growing set.  With summarization only in memory, performance still
degrades due to the cost of fetching all previous updates from the backing
store into the shim layer. Fetching the latest updates is essential for SC
operations. With summarization enabled on both disk and memory, latency does
not increase over time, and the implementation realizes throughput of 67K
operations/minute.

>>>>

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under a 1DC configuration. The YCSB workload was modified to issue 10
operations per transaction, with a default 50:50 read:write mix. Each operation
is assumed to be eventually consistent. NoTxn corresponds to a configuration
that does not use transactions. Compared to this, RC is only 12\% shower in
terms of latency with 512 clients, whereas RR is 2.3X slower. The difference
between RC and NoTxn is due to the meta-data overhead of recording transaction
information in the object state. For RR transactions, the cost of capturing and
maintaining a snapshot is the biggest source of overhead.

@@

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
configuration against a single replica (NoRep) and a strongly replicated
(StrongRep) \cf{1DC} deployment. The benchmark uses the default RUBiS bidding
mix, which has 15\% read-write interactions, which is representative of the
auction workload.  Without replication, NoRep trivially provides strong
consistency. However, this deployment does not scale beyond 1750 operations per
second. Strong replication offers better throughput at the cost of greater
latency due to inter-replica coordination. The \name deployment offers the
benefit of replication, while only paying the cost of coordination when
necessary.

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We use 128 clients and a single \name replica,
with all clients operating on the \emph{same} LWW register to stress test the
summarization mechanism. The shim layer cache (memory) is summarized every 64
updates, while the updates in the backing store (disk) are summarized every
4096 updates. Each point in the graph represents the average latency of the
previous 1000 operations. Each experiment is run for one minute.  Without
summarization, the average latency of operations increases exponentially to
almost one second, and only 13K operations were performed in a minute. Since
every operation has to reduce over the set of all previous operations,
operations take increasingly more time to complete since they must contend with
an ever growing set.  With summarization only in memory, performance still
degrades due to the cost of fetching all previous updates from the backing
store into the shim layer. Fetching the latest updates from the backing store
is essential for SC operations. With summarization enabled on both disk and
memory, latency does not increase over time, and the implementation realizes
throughput of 67K operations/minute.

### 2014-11-13 transactions.tex
edits

\subsection{Syntax and Semantics Extension}
\label{sec:syn_sem_ext}

@@

In order to illustrate the utility of declarative reasoning for transactions,
consider an extension of our running example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. The \name code for the
transactions is given below:
\vspace{-1em}

\noindent \begin{minipage}[t]{0.53\columnwidth}
\begin{codehaskell}
save amt =
  x <- dollar(classify $\cv_{sv}$)
  atomically x dollar do
    b <- withdraw c amt
    when b dollar deposit s amt
\end{codehaskell}
\end{minipage}
\begin{minipage}[t]{0.47\columnwidth}
\begin{codehaskell}
totalBalance =
  x <- dollar(classify $\cv_{tb}$)
  atomically x dollar do
    b1 <- getBalance s
    b2 <- getBalance s
    return b1 + b2
\end{codehaskell}
\end{minipage}

$\cv_{sv}$ and $\cv_{tb}$ are the contracts on the corresponding
transactions. The function \cf{classify} assigns the contracts
\emph{statically} to one of the transaction isolation levels offered by the
store; \cf{\$()} is meta-programming syntax for splicing the result into the
program. The \cf{atomically} construct invokes the enclosing operations at the
given isolation level $x$, ensuring that the effects of the operations are made
visible atomically. Our goal is to ensure that \cf{totalBalance} returns the
result obtained from a consistent snapshot of the object states.

@@

\noindent The key idea in the above definition is that the $\small \txnZ$ primitive
allows us to relate operations on different objects. The \cf{save} transaction
only needs to ensure that the two writes it performs are made visible
atomically. Since this is ensured by combining them in a transaction, \cf{save}
does not require any additional constraints, and $\psi_1$ is simply $\small
\true$.

@@

MAV semantics ensures that if some operation in a transaction $T_1$ witnesses
the effects of another transaction $T_2$, then subsequent operations in $T_1$
will also witness the effects of $T_2$. MAV semantics is useful for
maintaining the integrity of foreign key constraints, materialized views and
secondary updates. In order to implement MAV, a store only needs to keep track
of the set of transactions $S_t$ witnessed by the running transaction, and
before performing an operation at some replica, ensure that the replica
includes all the transactions in $S_t$. Hence, MAV is coordination-free. MAV
semantics is captured with the following contract:

>>>>

\subsection{Syntax and Semantics Extensions}
\label{sec:syn_sem_ext}

@@

In order to illustrate the utility of declarative reasoning for transactions,
consider an extension of our running example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. Our goal is to ensure
that \cf{totalBalance} returns the result obtained from a consistent snapshot
of the object states. The \name code for the transactions is given below:
\vspace{-1em}

\noindent \begin{minipage}[t]{0.53\columnwidth}
\begin{codehaskell}
save amt =
  x <- dollar(classify $\cv_{sv}$)
  atomically x dollar do
    b <- withdraw c amt
    when b dollar deposit s amt
\end{codehaskell}
\end{minipage}
\begin{minipage}[t]{0.47\columnwidth}
\begin{codehaskell}
totalBalance =
  x <- dollar(classify $\cv_{tb}$)
  atomically x dollar do
    b1 <- getBalance c
    b2 <- getBalance s
    return b1 + b2
\end{codehaskell}
\end{minipage}

$\cv_{sv}$ and $\cv_{tb}$ are the contracts on the corresponding
transactions. The function \cf{classify} assigns the contracts
\emph{statically} to one of the transaction isolation levels offered by the
store; \cf{\$()} is meta-programming syntax for splicing the result into the
program. The \cf{atomically} construct invokes the enclosing operations at the
given isolation level $x$, ensuring that the effects of the operations are made
visible atomically.

@@

\noindent The key idea in the above definition is that the $\small \txnZ$ primitive
allows us to relate operations on different objects.

The \cf{save} transaction only needs to ensure that the two writes it performs
are made visible atomically. Since this is ensured by combining them in a
transaction, \cf{save} does not require any additional constraints, and
$\psi_1$ is simply $\small \true$.

@@

MAV semantics ensures that if some operation in a transaction $T_1$ witnesses
the effects of another transaction $T_2$, then subsequent operations in $T_1$
will also witness the effects of $T_2$. MAV semantics is useful for maintaining
the integrity of foreign key constraints, materialized views and secondary
updates~\cite{BailisHAT}. In order to implement MAV, a store only needs to keep
track of the set of transactions $S_t$ witnessed by the running transaction,
and before performing an operation at some replica, ensure that the replica
includes all the transactions in $S_t$. Hence, MAV is coordination-free. MAV
semantics is captured with the following contract:
